~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project

### 2014-09-01 articles/2014/whisper/aricleEbauche/Draft_BriandBzeznikIrodsUserMeeting2014.tex
Minor corrections and added 2 schemas

\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
	Briand Xavier\footnote{Cnrs, Isterre, Whisper, Verce, email xav.briand@gmail.com}  & Bzeznik Bruno\footnote{CIMENT, Universit\'e Joseph Fourier, Technical header of CIMENT} \\
\end{tabular}}
\maketitle
%\thispagestyle{empty}
\subsection*{Abstract}
{\em
Like many projects in science of the universe, the seismological project Whisper is faced with massive data processing.
This leads to specific IT software for the project as well as suitable IT infrastructure. We present here both aspects.

@@

This is the partnership between these two IT components that has enabled a data-intensive processing and, also,  permits to the Whisper project to bring new scientific results.
}
~\\ \\ 
{\bf Keywords}:  Data-Intensive, Grid computing, Distributed storage, Seismic Noise, Whisper, CiGri, iRODS.

@@

Reflected by the Whisper project, we have to organize the science objectives with
the computer constraints. We have to take into account 
the duration of postdocs and PhD theses, as well as the availability of computer infrastructures and their ease of access. 
This leads to many questions about software development, including the genericity of computer code and 
the technical support. But it also influences in terms of choice of appropriate infrastrutcures.
%\vspace{-0.02cm}

Even if this project has his own ressources, such a problem of data-intensive requires specific tools able to 
organize distributed data management and acces to computational ressources: a data grid environment.
\vspace{-0.04cm}

@@

A part of the Whisper project is specificaly dedicated to the IT codes. This includes to design a specification, an implementation and some optimisations of the sequence of data-management as well as of the computations \footnote{see code-whisper.isterre.fr/html/ (part of the design)}.
This project uses its own IT resources (servers, dedicated bay) as well as the common IT infrastructure of the university.
We developed also adaptations for the IT infrastructures and we provide technical support for researchers.

@@

The first package provides a flexible way to process raw data, to specify a pipeline of pre-processing of signal.
The user starts by specifying a directory, a set of seismic stations and a set of dates. Then the codes scans the directory
and extracts all pieces of seismograms (also called traces) and rearranges them in a specific architecture of files in order to calculate the correlations to the next step.
We use here intensively the function 'read' of the library Obspy which allows to open most of the seismogram file formats.
The user also define his own sequence of processings. He can use the functions predefined but also the Python libraries he needs and,
moreover, he can add eventually his own functions.

@@

%Figure serveur CiGri et clusters, REST API

@@

%Figure JDL+Params-> campaign

@@

\begin{itemize}
  \item iRODS has a different transfert policy for "small" files (under 32MB). In this case, files are transiting through the server used by the client instead of a direct connection between the original resource and the client. In a high load context, with a lot of small file transfers, it leads to an overload of the concerned server. A possible solution to that problem could be to share the load among the servers of a given site. For that, we can imagine using a virtual ip address making a round robind on all the ip addresses of the servers of the site.
  \item We issued some network overloads: automatic staging, as set up for a site that has a lower bandwidth than the others may result in a lot of background iRODS processes doing the replication even when connection rate limit is set up on the servers and even when iget commands are aborted. Under some circumstances, it may completely overload a 1Gb/s ethernet link between 2 sites. Another overload encountered was when the user activates the -Q (use RBUDP (datagram) protocol for the data transfer) option. Even from a single node, this option can cause dramatic overload of network interfaces. Maybe iRODS should implement better rate limit control. 
  \item We made some preliminary tests of the python bindings of the iRODS API (called Pyrods). We noticed that when you have a lot of small operations to do from a python code (like in the Whisper project), either creating a lot of small files, or meta-data, Pyrods may be 10 times faster than a shell loop with i-commands inside. We are now making code examples for the other users to be able to take benefits from this method that would have been useful for the Whisper project.
\end{itemize}

>>>>

\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
	Briand Xavier\footnote{Cnrs, Isterre, Whisper, Verce, email xav.briand@gmail.com}  & Bzeznik Bruno\footnote{CIMENT, Universit\'e Joseph Fourier, Technical header of CIMENT} \\
\end{tabular}}
\maketitle
%\thispagestyle{empty}
\subsection*{Abstract}
{\em
Like many projects in science of the universe, the seismological project Whisper is faced with massive data processing.
This leads to specific IT softwares for the project as well as suitable IT infrastructures. We present here both aspects.

@@

This is the partnership between these two IT components that has enabled a data-intensive processing and, also,  permited the Whisper project to bring new scientific results.
}
~\\ \\ 
{\bf Keywords}:  Data-Intensive, Grid computing, Distributed storage, Seismic Noise, Whisper, CiGri, iRODS.

@@

Reflected by the Whisper project, we have to organize the science objectives with
the computer constraints. We have to take into account 
the duration of postdocs and PhD theses, as well as the availability of computer infrastructures and their ease of access. 
This leads to many questions about software development, including the genericity of computer code and 
the technical support. But it also influences in terms of choice of appropriate infrastructures.
%\vspace{-0.02cm}

Even if this project has its own ressources, such a problem of data-intensive processing requires specific tools able to 
organize distributed data management and access to computational ressources: a data grid environment.
\vspace{-0.04cm}

@@

A part of the Whisper project is specificaly dedicated to the IT codes. This includes the design of a specification, an implementation and some optimisations of the sequence of data-management as well as of the computations \footnote{see code-whisper.isterre.fr/html/ (part of the design)}.
This project uses its own IT resources (servers, dedicated bay) as well as the common IT infrastructure of the university.
We developed also adaptations for the IT infrastructures and we provide technical support for researchers.

@@

The first package provides a flexible way to process raw data, to specify a pipeline of pre-processing of signal.
The user starts by specifying a directory, a set of seismic stations and a set of dates. Then the codes scans the directory
and extracts all pieces of seismograms (also called traces) and rearranges them in a specific architecture of files in order to calculate the correlations to the next step.
We use here intensively the function 'read' of the library Obspy which allows to open most of the seismogram file formats.
The user also defines his own sequence of processings. He can use the functions predefined but also the Python libraries he needs and,
moreover, he can add eventually his own functions.

@@

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaCigriArch} Cigri communications}
\includegraphics[width=8cm]{cigri_arch.png}
\end{center}

@@

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaCigriJDL} Cigri jobs campaign submission}
\includegraphics[width=8cm]{cigri_sub.png}
\end{center}

@@

\begin{itemize}
  \item iRODS has a different transfert policy for "small" files (under 32MB). In this case, files are transiting through the server used by the client instead of a direct connection between the original resource and the client. In a high load context, with a lot of small file transfers, it leads to an overload of the concerned server. A possible solution to that problem could be to share the load among the servers of a given site. For that, we can imagine using a virtual ip address making a round robind on all the ip addresses of the servers of the site.
  \item We issued some network overloads: automatic staging, as set up for a site that has a lower bandwidth than the others may result in a lot of background iRODS processes doing the replication even when connection rate limit is set up on the servers and even when iget commands are aborted. Under some circumstances, it may completely overload a 1Gb/s ethernet link between 2 sites. Another overload encountered was when the user activates the -Q (use RBUDP (datagram) protocol for the data transfer) option. Even from a single node, this option can cause dramatic overload of network interfaces. Maybe iRODS should implement better rate limit control. 
  \item We made some preliminary tests of the python bindings of the iRODS API (called Pyrods). We noticed that when you have a lot of small operations to do from a python code (like in the Whisper project), either creating a lot of small files, or meta-data, Pyrods may be 10 times faster than a shell loop with i-commands inside. We are now making code examples for the other users to be able to take benefits from this method that would have been useful for the Whisper project.
  \item We also noticed that a multiprocess program made with python and the PyRODS API improves the speed of a recursive transfer of a collection containing a lot of small files (thousands of 64kb files for example). We can conclude that iget/iput commands are not optimized for that kind of transfers.
\end{itemize}
